---
title: "Math 3636 Final Project Group 9"
output:
  html_notebook: default
  pdf_document: default
---

## Loading Data Into R

```{r}
setwd("D:\\Documents\\R Codes\\Final Project")
```


```{r}
heart <- read.csv("HeartData.csv")

heart$sex <- factor(heart$sex)
heart$cp <- factor(heart$cp)
heart$fbs <- factor(heart$fbs)
heart$restecg <- factor(heart$restecg)
heart$exang <- factor(heart$exang) 
heart$slope <- factor(heart$slope)
heart$ca <- factor(heart$ca)
heart$thal <- factor(heart$thal)

summary(heart)

```

## Outliers:
We checked for outliers within our data by visualualizing the data. 
```{r}
dev.new(width = 4, height = 4)

plot(age ~ condition, data = heart)

AgeOutliersRange <- c((quantile(heart$age)[2] - (quantile(heart$age)[4] - quantile(heart$age)[2])*1.5), (quantile(heart$age)[4] + (quantile(heart$age)[4] - quantile(heart$age)[2])*1.5))

summary(heart$age)
AgeOutliersRange

AgeOutInd <- which(heart$age < AgeOutliersRange[1] | heart$age > AgeOutliersRange[2])

#heart <- heart[- AgeOutInd, ]

#plot(age ~ condition, data = heart)


plot(trestbps ~ condition, data = heart)

TrestbpsOutliersRange <- c((quantile(heart$trestbps)[2]-(quantile(heart$trestbps)[4] - quantile(heart$trestbps)[2])*(1.5)), (quantile(heart$trestbps)[4] + (quantile(heart$trestbps)[4] - quantile(heart$trestbps)[2])*(1.5)))

summary(heart$trestbps)
TrestbpsOutliersRange

TrestbpsOutInd <- which(heart$trestbps < TrestbpsOutliersRange[1] | heart$trestbps > TrestbpsOutliersRange[2])

heart <- heart[- TrestbpsOutInd, ]

plot(trestbps ~ condition, data = heart)


plot(chol ~ condition, data = heart)

CholOutliersRange <- c((quantile(heart$chol)[2] - (quantile(heart$chol)[4] - quantile(heart$chol)[2])*1.5), (quantile(heart$chol)[4] + (quantile(heart$chol)[4] - quantile(heart$chol)[2])*1.5))

summary(heart$chol)
CholOutliersRange

CholOutInd <- which(heart$chol < CholOutliersRange[1] | heart$chol > CholOutliersRange[2])

heart <- heart[- CholOutInd, ]

plot(chol ~ condition, data = heart)


plot(thalach ~ condition, data = heart)

ThalachOutliersRange <- c((quantile(heart$thalach)[2] - (quantile(heart$thalach)[4] - quantile(heart$thalach)[2])*1.5), (quantile(heart$thalach)[4] + (quantile(heart$thalach)[4] - quantile(heart$thalach)[2])*1.5))

summary(heart$thalach)
ThalachOutliersRange

ThalachOutliersInd <- which(heart$thalach < ThalachOutliersRange[1] | heart$thalach > ThalachOutliersRange[2])

heart <- heart[- ThalachOutliersInd, ]

plot(thalach ~ condition, data = heart)


plot(oldpeak ~ condition, data = heart)

OldpeakOutliersRange <- c((quantile(heart$oldpeak)[2] - (quantile(heart$oldpeak)[4] - quantile(heart$oldpeak)[2])*1.5), (quantile(heart$oldpeak)[4] + (quantile(heart$oldpeak)[4] - quantile(heart$oldpeak)[2])*1.5))

summary(heart$oldpeak)
OldpeakOutliersRange

OldpeakOutliersInd <- which(heart$oldpeak < OldpeakOutliersRange[1] | heart$oldpeak > OldpeakOutliersRange[2])

heart <- heart[- OldpeakOutliersInd, ]

plot(oldpeak ~ condition, data = heart)
```


## Confusion Matrix for sex and condition:
```{r}
table(heart$sex,heart$condition)
```


## Linear Model:

```{r}
heartfit1 <- lm(condition ~., data = heart)

summary(heartfit1)

```

Plot residuals of linear model:

```{r}
par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(heartfit1$fitted.values, ylab = "Predicted Values")
plot(heartfit1$residuals, ylab = "Residuals")

plot(heartfit1)
```


## Linear Model Reduced:

```{r}
heartfit1Reduced <- lm(condition ~sex + cp + ca + thal + trestbps + chol + slope + oldpeak, data = heart)

summary(heartfit1Reduced)

```

Plot residuals of linear model Reduced:

```{r}
par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(heartfit1Reduced$fitted.values, ylab = "Predicted Values")
plot(heartfit1Reduced$residuals, ylab = "Residuals")

plot(heartfit1Reduced)
```


## Generalized Linear Model probit:

```{r}
heartfitprobit <- glm(formula = condition~., family = binomial(link = "probit"), data = heart)

summary(heartfitprobit)
```

Plot residuals of this model:

```{r}
par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(heartfitprobit$fitted.values, ylab = "Predicted Values")
plot(heartfitprobit$residuals, ylab = "Residuals")

plot(heartfitprobit)
```
This model is good because Residuals are uniform.

## Generalized Linear Model Probit Reduced:

```{r}
heartfitprobitreduced <- glm(formula = condition ~ sex + cp + ca + thal + trestbps + chol + slope + oldpeak, family = binomial(link = "probit"), data = heart)

summary(heartfitprobitreduced)
```

Plot residuals of this model:

```{r}
par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(heartfitprobitreduced$fitted.values, ylab = "Predicted Values")
plot(heartfitprobitreduced$residuals, ylab = "Residuals")

plot(heartfitprobitreduced)
```
This model is good because Residuals are uniform.


## Generalized Linear Model Logit:

```{r}
heartfit2 <- glm(formula = condition ~., family = binomial(link = "logit"), data = heart)

summary(heartfit2)

```

Plot residuals of this model:

```{r}
par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(heartfit2$fitted.values, ylab = "Predicted Values")
plot(heartfit2$residuals, ylab = "Residuals")

plot(heartfit2)
```
This model is good because Residuals are uniform.


## Generalized Linear Model Logit Reduced:

```{r}
heartfit2reduced <- glm(formula = condition ~ sex + cp + ca + thal + trestbps + chol + slope + oldpeak, family = binomial(link = "logit"), data = heart)

summary(heartfit2reduced)

```

Plot residuals of this model:

```{r}
par(mar=c(4,4,1,1), mfrow=c(1,2))
plot(heartfit2reduced$fitted.values, ylab = "Predicted Values")
plot(heartfit2reduced$residuals, ylab = "Residuals")

plot(heartfit2reduced)
```
This model is good because Residuals are uniform.


## Predict Linear Model and Calculate Accuracy:

```{r}
pred <- predict(heartfit1, newdata = heart, type = "response")

pred2 <- ifelse(pred < 0.54, 0, 1)

res <- data.frame(Obs = heart$condition, Pre = pred2)

calACC <- function(p1, p2) {
  
  return(sum(diag(table(p1, p2))) / length(p1))

}

calACC(p1 = heart$condition, p2 = pred2)

```

Confusion Matrix:

```{r}
table(heart$condition, pred2)
```


## Predict Linear Model Reduced and Calculate Accuracy:

```{r}
pred <- predict(heartfit1Reduced, newdata = heart, type = "response")

pred2 <- ifelse(pred < 0.54, 0, 1)

res <- data.frame(Obs = heart$condition, Pre = pred2)

calACC <- function(p1, p2) {
  
  return(sum(diag(table(p1, p2))) / length(p1))

}

calACC(p1 = heart$condition, p2 = pred2)

```

Confusion Matrix:

```{r}
table(heart$condition, pred2)
```


## Predict Generalized Linear Model Probit and Calculate Accuracy:

```{r}
pred <- predict(heartfitprobit, newdata = heart, type = "response")

pred2 <- ifelse(pred < 0.54, 0, 1)

res <- data.frame(Obs = heart$condition, Pre = pred2)

calACC <- function(p1, p2) {
  
  return(sum(diag(table(p1, p2))) / length(p1))

}

calACC(p1 = heart$condition, p2 = pred2)

```

Confusion Matrix:

```{r}
table(heart$condition, pred2)
```


## Predict Generalized Linear Model Reduced Probit and Calculate Accuracy:

```{r}
pred <- predict(heartfitprobitreduced, newdata = heart, type = "response")

pred2 <- ifelse(pred < 0.54, 0, 1)

res <- data.frame(Obs = heart$condition, Pre = pred2)

calACC <- function(p1, p2) {
  
  return(sum(diag(table(p1, p2))) / length(p1))

}

calACC(p1 = heart$condition, p2 = pred2)

```

Confusion Matrix:

```{r}
table(heart$condition, pred2)
```


## Predict Generalized Linear Model Logit and Calculate Accuracy:

```{r}
pred <- predict(heartfit2, newdata = heart, type = "response")

pred2 <- ifelse(pred < 0.54, 0, 1)

res <- data.frame(Obs = heart$condition, Pre = pred2)

calACC <- function(p1, p2) {
  
  return(sum(diag(table(p1, p2))) / length(p1))

}

calACC(p1 = heart$condition, p2 = pred2)

```

Confusion Matrix:

```{r}
table(heart$condition, pred2)
```


## Predict Generalized Linear Model Reduced Logit and Calculate Accuracy:

```{r}
pred <- predict(heartfit2reduced, newdata = heart, type = "response")

pred2 <- ifelse(pred < 0.54, 0, 1)

res <- data.frame(Obs = heart$condition, Pre = pred2)

calACC <- function(p1, p2) {
  
  return(sum(diag(table(p1, p2))) / length(p1))

}

calACC(p1 = heart$condition, p2 = pred2)

```

Confusion Matrix:

```{r}
table(heart$condition, pred2)
```

```{r}
table(heart$sex, heart$condition)
```

